---
title: "Project 1 - DATS 6101"
author: "Randomly-Generated: Tran Hieu Le, Fahim Ishrak, Zhilin Wang, Totyana Hill "
date: "9/29/2019"
output:
   html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE, warning = FALSE)
options(scipen=9, digits = 3) 
```

```{r define_loadPkg}
loadPkg = function(x) { if (!require(x,character.only=T, quietly =F)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T, quietly=T)) stop("Package not found") } }
```

### Chapter 1: Introduction


### Chapter 2: Description of Data

#### 2.1 Source Data

The movie dataset for our project is taken from Kaggle (link: https://www.kaggle.com/danielgrijalvas/movies).

This data was scraped from The Internet Movie Database (IMDb) website (https://www.imdb.com/) using Python scripts before being uploaded to Kaggle as a CSV file.

Limitation of this dataset: 
* Because the source's owner scraped the dataset from a website, there may be a percentage of data that were not successfully retrieved.
* IMDb does not have enough information in countries where the movie industry is less developed. Therefore, we do not actually study the global movie industry. Instead, we are doing research on most popular movies.
* Since the reviews of IMDb comes solely from audiences, the score of movies is neither objective nor well-evaluated. The fanboys of the movie or its actors/ actress will likely grade the movie 10 score, whereas the haters often give a rating of one. 
 
#### 2.2 Data Explanation

The CSV contains records of 6820 movies in about three decades (1986-2016). There are a lot of information about movie included in the CSV file.

The following is the full list of movie attributes in the columns with their explanation:

*	budget: the amount of money was spent to produce a movie
*	company: the movie's producer
*	country: the country where the movie is produced
*	director: the director of the movie
*	genre: the main genre of the movie
*	gross: the revenue of the movie
*	name: the name of the movie
*	rating: a system determines the suitability of movies for different audiences (classified by ages)
*	released: the release date (YYYY-MM-DD)
*	runtime: the duration of the movie
*	score: IMDB user rating
*	votes: the number of user votes
*	star: the main actor/actress
*	writer: the writer of the movie
*	year: the year of release

(P/s: The raw data will be shown in our data cleaning process to provide an overview of the CSV file)

To be more detailed about rating, we will introduce the Motion Picture Association of America (MPAA) film rating system:

*	G – General Audiences: All ages admitted. Nothing that would offend parents for viewing by children.
*	PG – Parental Guidance Suggested: Some material may not be suitable for children. Parents urged to give “parental guidance”. May contain some material parents might not like for their young children.
*	PG-13 – Parents Strongly Cautioned: Some material may be inappropriate for children under 13. Parents are urged to be cautious. Some material may be inappropriate for pre-teenagers.
*	R – Restricted: Under 17 requires accompanying parent or adult guardian. Contains some adult material. Parents are urged to learn more about the film before taking their young children with them.
*	NC-17 – Adults Only: No One 17 and Under Admitted. Clearly adult. Children are not admitted.

Any movies which are not assigned the final rating by MPAA are uncut/extended versions with different contents from their theatrical releases, or commercial products used to advertise for the main television or DVDs series. Some movies with local labels are published to serve audiences in a certain country or region, hence, they are not submitted for a rating.

Since our goal is to research the theatrical exhibition worldwide, we will remove all movies without MPAA rating. Furthermore, any movie attributes that do not support our analysis will be excluded to ensure our dataset is clean and helpful. This process will be conducted through our data cleaning in next part. 

 

#### 2.3 Data Cleaning

Step 1: Load dataset from csv file and save it as movies.

```{r load_data, include = T}

loadPkg("ggplot2")
loadPkg("dplyr")

# read dataset
movies <- read.csv("movies.csv")
str(movies)

```


Step 2: Clean the dataframe. We will save our new dataframe as movies_new

*	At first, we will remove duplicated data. Fortunately, all the data are distinct.

```{r duplicates}
# remove duplicates if they exist
movies <- unique(movies)

```

*	Secondly, we will remove columns that will not be used in our analysis: company, director, released, runtime, star and writer.

```{r rmcol}

#remove unused columns and save the subset as movies_new 

movies_new <- subset(movies, select = - c(company,director,released,runtime,star,writer))

```


*	As our statement in the data explanation, our analysis concentrates on global theatrical movies so we will exclude movies that are uncut/extended versions, TV/DvDs advertisements or local products. 

```{r ratings_rework}

# filter for movies with MPAA film rating system
movies_new <- subset(movies_new, movies_new$rating%in%c("G","NC-17","PG-13","R","PG"))


# apply factor again to new dataset
movies_new$rating <- factor(movies_new$rating) 
movies_new$genre <- factor(movies_new$genre)
movies_new$country <- factor(movies_new$country)


```


*	There are some movies with budget = 0, meaning that their budget data were not recorded. To treat these values, we will change the budget from 0 to NA. We use the same treatment with gross data. Fortunately, there are no unrecorded gross data. 

```{r NAs}

# assign NA to movies with budget = 0 
movies_new$budget <- replace(movies_new$budget, movies_new$budget==0, NA)

# assign NA to movies with gross = 0. Fortunately, we do not have any unrecorded gross data
#movies_new$gross <- replace(movies_new$gross, movies_new$gross==0, NA)

```


*	Finally, since we will do research on movie commerce, we will create two new columns to support our analysis: Profit and Return on Investment (ROI).

```{r createcol}

movies_new$profit <- movies_new$gross - movies_new$budget # create profit column

movies_new$roi <- (movies_new$profit)*100/(movies_new$budget) # create ROI column


```



Step 3: Show our new dataframe: movies_new

```{r datastructure, include = T}
# show dataframe structure
str(movies_new)


```

#### 2.3 Data Visualization

```{r datahist, include = T}

ggplot(data = movies_new, aes(x = year)) +
  geom_histogram(alpha = 0.5, fill = "blue", col = "red", binwidth = 1) +
  ggtitle("Histogram of published movies from 1986-2016") +
  xlab("Year") +
  ylab("Number of Movies") +
  theme(plot.title = element_text(hjust=0.5))
  
```

* Our dataset has over 200 movies per year and it is pretty good for our analysis. We do not have to treat any years with significant number of movies.


#### 2.4 Movie industry 

In this part, we will have an overall visualization about the global movie industry and figure out which country is the largest movie industry. 

* Let's see the which countries have the highest number of published movies.

```{r country, include = T}

# Name the country with max counts
name_max <- names(which.max(table(movies_new$country)))

# percentage that country accounts for
pp_max <- nrow(subset(movies_new, movies_new$country == name_max))*100 / nrow(movies_new)

country_table <- table(movies_new$country) # create a table with country and count no. of movies
country_table <- data.frame(country_table) # save table into a dataframe
country_table <- subset(country_table, Freq > 20) # select countries with more than 20 movies in total

# bar plot for no. of movies by country
ggplot(country_table, aes(x = reorder(Var1, Freq), y = Freq)) + 
  geom_bar(stat = 'identity', fill ='orange') + coord_flip() +
  ylab("Number of movies") +
  xlab("Country") +
  ggtitle(("Number of movies by country")) +
  theme(plot.title = element_text(hjust = 0.5)) 

```

* A geographic visualization of the global movie industry. 

P/s: Countries and locations labeled with grey color have few global products or their movies are not popular in the world. Their movie industries primarily serve their local citizens.

```{r countrymap, include = T}


loadPkg("maps")
loadPkg("viridis")

colnames(country_table)[1] <- c("region") # name column as region

# prepare label coordinate for USA
USA_map <- map_data("world", region = c("USA"))
USA_name <- USA_map %>% group_by(region) %>% summarise(long = mean(long), lat = mean(lat))

world_map <- map_data("world")
movie_map <- left_join(world_map, country_table, by = "region" ) # join movie dataset to map dataset by country/region

ggplot(data = movie_map, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = Freq), col = "grey") +
  geom_text(aes(label = region), data = USA_name,  size = 3, hjust = -0.7, vjust = 2.5) +
  scale_fill_viridis_c(option = "plasma") +
  ggtitle("Global Movie Industry") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Longitude") + ylab("Latitude") 
  
```



* The country with the highest number of published movies over the three decades is `r name_max`.
* `r name_max` accounts for approximately `r round(pp_max)` percentage of total number of movies.

It is easy to realize that the biggest movie industries in the world are also the most developed countries with strong economic potential.

* USA is the largest movie industry, as expected from the biggest economy in the world. Their advance in technology also makes a considerable contribution to the growth of their movie industry.
* In Europe, the biggest movie industries are also the most developed countries such as United Kingdom, France and Germany.
* In Asia, China, Japan and Australia are the most growing movie industries. 

Now, it is interesting to investigate if the largest movie industry can reflect the global movie industry. To do this, we will conduct a Chi-squared test Goodness of Fit:


```{r GoodnessOfFit, include = T}

# save all genres into a vector
fullgenrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Family","Fantasy","Horror","Musical","Mystery","Romance", "Sci-Fi","Thriller", "War","Western")

USA_movies <- subset(movies_new, movies_new$country == "USA")

exp_freq <- c("")
USA_freq <- c("")

# calculate expected frequencies
j = 1
for(i in fullgenrelist) {exp_freq[j] <- nrow(subset(movies_new, movies_new$genre == i))/ nrow(movies_new); j = j + 1}
exp_freq <- as.numeric(as.character(exp_freq))

# calculate observed frequencies
k = 1
for(i in fullgenrelist) {USA_freq[k] <- nrow(subset(USA_movies, USA_movies$genre == i)); k = k + 1}
USA_freq <- as.numeric(as.character(USA_freq))

# Chi-test for Goodness of Fit
chiGoF <- chisq.test(USA_freq, p = exp_freq)


# results
chiGoF



```

* Null Hypothesis: The USA data distribution matches the global dataset.
* The p-value of the test is `r format(chiGoF$p.value, digits = 3)` which is much smaller than significance level 0.05.
* We reject the null and conclude that the USA data distribution does not match the whole dataset at 0.05 level.
* Unfortunately, we do not have enough evidence to use USA as a sample to study the global movie industry.


### Chapter 3: Movie categories

In this section, we will study about the movie categories and figure out which genres are the most popular. We will also examine if movie genre and votes have any relationship and discover the most favorite genre based on the number of votes per movie (we define the number of votes as the basis of favorite).

Since the genre of a movie is determined by its content (adult scenes, level of violence, romance elements, criminal activities), it is argued that there is a relationship between movie genre and movie rating, i.e. movie genre is a factor to define the rating. We will examine this argument by hypothesis test.  



#### 3.1 SMART Question
##### Does the movie genre have any effects on the movie rating, and which is the most favorite genre?

#### 3.2 Number of published movies by genres

* We will calculate the number of published movies by genre in 1986-2016.

```{r count, include = T}

# show the number of published movies by genre
table(movies_new$genre)

# save new dataset including 8 popular genres
movies_fam <- subset(movies_new, movies_new$genre%in%c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror"))
movies_fam$genre <- factor(movies_fam$genre)

```

* Below is the visualization of the number of movies by genre.

```{r pltcount, include =T}

ggplot(movies_new, aes(x=genre, fill = genre)) +
  geom_bar(stat="count", ) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("The number of movies published by genre from 1986 to 2016") +
  guides(fill = guide_legend(title="Genre"))


```

* As we see from the table and the bar charts, Action, Adventure, Animation, Biography, Comedy, Crime, Drama, and Horror are the most popular genres. Action, Comedy and Drama have most products in the three decades.

* In contrast, Family, Fantasy, Musical, Mystery, Romance, Sci-Fi, Thriller, War and Western movies seem not to be preferred by film companies and directors. 

* Indeed, the elements of Family, Fantasy, Musical, Mystery, Romance, Sci-Fi, Thriller also exist in other genres and this inclusion improved the quality of Adventure, Animation, Biography, Comedy, Crime, Drama, and Horror movies. However, when a single of these elements is attributed to the main content of a movie, it makes audiences feel monotonous and boring.

* War and Western movies seem not to interest international viewers. The reason may be that they focus a lot on historical events and stories which appear frequently in papers, books and documentaries. Their main contents on the history bring little entertainment to audiences.

* In the next parts, we will focus on most popular genres: Action, Adventure, Animation, Biography, Comedy, Crime, Drama, and Horror.


```{r save_a_genrelist}

# save 8 most popular genres into a vector
genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")


```


#### 3.3 Genre and rating relationship


``` {r contable}

# create contigency table with all genres

# contable1 = table(movies_new$rating, movies_new$genre)
# contable1

# create contigency table with popular genres
contable2 <- table(movies_fam$rating, movies_fam$genre)
contable2

````

* To test the dependence of genre and rating, we will conduct a Chi-squared Test Independence to examine the null hypothesis: Movie genre and movie rating are independent.


```{r chitest, include = T}

# chi-square test with all genres
#chitest1 <- chisq.test(contable1)

# results
#chitest1
#chitest1$statistic
#chitest1$parameter
#chitest1$expected
#chitest1$p.value


# chi-square test with popular genres
chitest2 <- chisq.test(contable2)

# results
chitest2
#chitest2$statistic
#chitest2$parameter
#chitest2$expected
#chitest2$p.value



```


* The p-value in the chi-squared test is `r format(chitest2$p.value, digits = 3)`, much smaller than the significance level 0.05. We reject the null.
* Conclusion: The movie genre and rating are NOT independent.
* Therefore, there is an evidence that movie genre affects a movie's rating.


#### 3.4 Vote and Genre relationship

* It is obvious that some genres are of great interest to audiences. This may affect users' votes for a movie since they will always suggest their favorite genre. We will identify whether the genre of a movie influences users’ voting.
* To do this we will conduct an anova test to see if different genres perform similarly in users' votes.

```{r Vote&Genre, include = T}
anova_vg <- aov(votes ~ genre, movies_fam )
#summary(anova_vg)
tukey_vg <- TukeyHSD(anova_vg)
tukey_vg

```

* From our anova test, the p-value is `r format(summary(anova_vg)[[1]][["Pr(>F)"]][1], digit = 3)` which is much smaller than significance level 0.05. We reject the null hypothesis that the voting averages of different genres are all the same.

* Since the voting averages by genre are not all the same, there is an evidence that movie genre has some relationship with users' votes. There are some certain genres that greatly interest audiences, hence, they receive more votes than others.

* From the Tukey test, we can divide 8 genres into two groups. Each pair of genres within a group are not significant but two genres from two different groups will have significant means.

Group 1: Action, Adventure, Animation
Group 2: Biography, Comedy, Horror, Drama
Crime is an exception since Crime-Adventure and Crime-Biography are not significant.

* Prediction: One of the two groups will contain the most favorite genres, while the least favorite genres will be in the other. Crime seems to stay in the middle of the two groups.


#### 3.4 Most favorite genres

* We will confirm our prediction in the above part and figure out the most favorite group. 
* We will divide the timeline into three decades: 1986-1995, 1996-2005, 2006-2016 to see how audiences vote in different periods.

``` {r databyperoid}

# create subsets representing different decades
movies1 <- subset(movies_fam, movies_fam$year%in%c(1986:1995)) 
movies2 <- subset(movies_fam, movies_fam$year%in%c(1996:2005))
movies3 <- subset(movies_fam, movies_fam$year%in%c(2006:2016)) 
#str(movies1)
#str(movies2)
#str(movies3)
```


* We determine the most favorite genre based on the number of votes per movie. Below is the ratio of average votes by genre in  different periods.

```{r def}

# define a function to plot
pltm = function(x) {
  genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")
  vote_avg <- c("") # create an empty vector that we will use to save average of votes by genre later
  j = 1
  
  #calculate the average of votes by genre and save them into our vector
  for(i in genrelist) {vote_avg[j] <- mean((subset(x, x$genre == i))$votes, na.rm = T); j = j + 1} 
  
  df = data.frame(meanvote = vote_avg, genrevote = genrelist) # combine two vectors (average of votes and genre names) into a dataframe
  df$meanvote <- as.numeric(as.character(df$meanvote)) # change the value from factor to numeric
  
  # plot a pie chart to see the ratio of average votes by genre
  p <- ggplot(df, aes(x="", y=meanvote, fill = genrevote)) +
    geom_bar(width=1, stat = "identity", color = "white") +
    coord_polar("y", start=0) +
    guides(fill=guide_legend(title="Genre")) +
    theme(axis.text  = element_blank(), axis.ticks = element_blank(), axis.title = element_blank()) +
    geom_text(aes(label = paste0( scales::percent(meanvote / sum(meanvote)))), position = position_stack(vjust = 0.5))
  

  return(p)


}
```

* A graph of average votes in each year:

```{r votechange}

Genre_vote <- subset(movies_fam, select = c(genre, votes, year))
Genre_vote <- Genre_vote %>% group_by(year, genre) %>% summarise(votes = mean(votes))

 
ggplot(data = Genre_vote, aes(x = year, y = votes, col = genre)) +
  geom_line() +
  geom_point() +
  ggtitle("Average votes by genre in 1986-2016") +
  theme(plot.title = element_text(hjust = 0.5)) 



```

* Ratio of average votes by genre:

```{r genrevotes, echo = F}

# No. of votes per movie in the whole period
pltm(movies_fam) +
  ggtitle("Ratio of average votes in 1986-2016") +
  theme(plot.title = element_text(hjust = 0.5))

# No. of votes per movie in 1986-1995
pltm(movies1) +
  ggtitle("Ratio of average votes in 1986-1995") +
  theme(plot.title = element_text(hjust = 0.5)) 
 

# No. of votes per movie in 1996-2005
pltm(movies2) +
  ggtitle("Ratio of average votes in 1996-2005") +
  theme(plot.title = element_text(hjust = 0.5))

# No. of votes per movie in 2006-2016
pltm(movies3) +
  ggtitle("Ratio of average votes in 2006-2016") +
  theme(plot.title = element_text(hjust = 0.5))


```

* As we see from the plots, Action, Animation and Adventure which belong to group 1 always receive more votes than Biography, Drama, Comedy and Horror which belong to group 2. Crime is always in the middle of the two groups except the first decade. This observation explained our prediction from previous part. 

* During the first period, Crime were the most favorite genre and Animation placed in the second. In the next periods, Action and Adventure grew as the most favorite genres and still maintained their posistions until 2016. In contrast, Animation and Crime dropped to third and fourth places after the first 10 years, respectively.

* Starting as the least favorite genre in the top 4, Action became the most favorite type of movie. The first pie chart also shows that Action is the most favorite genre in three decades. The growth of Action movies may be the result of the development of technology. The appearance of modern technologies in recent years has enhanced the visual effects, sounds and camera views ... And Action movies seem to take huge advatages from these improvements compare to other genres. Additionally, the groundbreaking ideas and diverse contents in Action movies seem to be of great interest to people.


### Chapter 4: Movie Scores

As we stated in the limitation of our dataset, the fanboys will likely grade the movie 10 score, while the haters often give a rating of one. This will significantly skew the movie score with either huge fanboys or huge haters. If we assume that the number of votes depend on fanboys and haters. We expect to see that movies with a lot of votes will score well while movies with few votes will receive bad grades.

In this section, we will research the relationship between user votes and the score to examine our prediction. Furthermore, we also build a hypothesis to test whether movie genre is a factor to the score. 

#### 4.1 SMART Question
##### Is there any relationships between the votes and the scores, and are the score means the same across all genres?

#### 4.2 Scores and Votes Relationship

We will divide votes and scores into different categories by our own criteria.

* Votes:

- High votes: from 1,000,000  
- Medium votes: from 500,000 to below 1,000,000
- Low votes: below 500,000

* Scores:

- Good score: from 7.5
- Medium score: from 5.0 to below 7.5
- Bad score: below 5.0

Now, we plot the distribution of votings across the scores using boxplots and mapping plots.

```{r pltVotesVsScore, include = T}

high_votes <- subset(movies_fam, votes > 1e+6 | votes == 1e+6)
medium_votes <- subset(movies_fam, votes < 1e+6 & votes > 500000 | votes == 500000)
low_votes <- subset(movies_fam, votes < 500000)


# boxplot for score distribution by different ranges of votes
ggplot() +
  ggtitle("Score distribution with different groups of votes") +
  geom_boxplot(data=high_votes, aes(x=1,y=score,fill="1")) +
  geom_boxplot(data=medium_votes, aes(x=2,y=score,fill="2")) +
  geom_boxplot(data=low_votes, aes(x=3,y=score,fill="3")) +
  theme(plot.title = element_text(hjust=0.5),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill=guide_legend(title = "Vote categories")) +
  scale_fill_manual(labels=c("High","Medium","Low"),values = c("blue","red","yellow"))
  
# geomplot for score and vote
ggplot() +
  ggtitle("Votes distribution over score") +
  geom_point(data=movies_fam,mapping=aes(x=score,y=votes,col=genre)) +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_hline(linetype="dashed", aes(yintercept = 1e+6)) + 
  geom_vline(linetype="dashed", aes(xintercept = 7.5)) +
  geom_hline(linetype="dashed", aes(yintercept = 500000)) + 
  geom_vline(linetype="dashed", aes(xintercept = 5.0)) +
  annotate("text", x = 8.6, y = 15*(1e+5), label = "High Votes\nHigh Score") +
  #annotate("text", x = 6.0, y = 15*(1e+5), label = "High Votes\nMedium Score") +
  annotate("text", x = 2.5, y = 15*(1e+5), label = "High Votes\nLow Score") +
  #annotate("text", x = 8.6, y = 750000, label = "Medium Votes\nHigh Score") +
  #annotate("text", x = 6.0, y = 750000, label = "Medium Votes\nMedium Score") +
  #annotate("text", x = 2.5, y = 750000, label = "Medium Votes\nLow Score") +
  annotate("text", x = 8.6, y = 250000, label = "Low Votes\nHigh Score") +
  #annotate("text", x = 6.0, y = 250000, label = "Low Votes\nMedium Score") +
  annotate("text", x = 2.5, y = 250000, label = "Low Votes\nLow Score")

#qqnorm(movies_new$votes)
#qqnorm(movies_new$score)

```

* As shown in boxplots, the movies with high votes always score well, while the movies with bad scores always have very low votes. 
* However, high score movies do not certainly have high votes and low votes movies do not mean that they will earn bad scores. 
* The mapping plot also gives us the same observation, movies with votes over 1 million always get high scores. And at around 500000 votes and below, the score varies extensively. Movies with scores below 5.0 have low votes.
* It seems that our prediction is not totally correct. The visualization matches our expectation for most favorite movies (with high votes) but in the cases of least voted movies the scores are not certainly bad. Our assumption seems to be inaccurate for low values of votes. The reason is that there are many good movies which belong to least favorite group (Biography, Crime, Horror and Drama). Therefore, these movies have a small number of viewers which makes them earn few votes despite their high grades.



#### 4.3 Score by genres and ratings

##### By genres:

* Anova test score means by famous genres: 

```{r AOV_score&genre, include =T}

anova_sg <- aov(score ~ genre, movies_fam)
summary(anova_sg)

```

* Null Hypothesis: The score means of all genres are the same.
* The p-value of the test is `r format(summary(anova_sg)[[1]][["Pr(>F)"]][1], digits = 3)`, which is very close to 0 and much smaller than significance level 0.05.
* We reject the null: the average scores of different genres are NOT all the same.
* It seems that movie genre has some influences on the score. It is understandable since people tend to give good scores for their favorite genres.


```{r pltScore&Genre, include = T}
ggplot() +
  geom_boxplot(data=movies_fam, aes(x=genre, y=score, fill=genre)) +
  ggtitle("Score distribution by genres") +
  theme(plot.title=element_text(hjust=0.5))
  
```

* As shown in the boxplots, the following pairs have pretty the same tall boxplots and overlapped interquartile ranges: Action-Comedy, Animation-Crime, Animation-Drama and Crime-Drama.
* This illustrates that these pairs seem to be NOT significant in their mean values. 
* We will conduct an ad hoc procedure using Tukey test to see which pairs are significant and check whether our visualization is correct. 

```{r TukeyTest1, include = T}

tukey1 <- TukeyHSD(anova_sg)
tukey1
```


* As we see from our Tukey test, there are no significant differences in means of the following pairs: Action-Comedy, Animation-Crime, Animation-Drama, Crime-Drama, while the other pairs are significant.
* The results match our observation from the boxplots.




### Chapter 5: Movie Commerce

In this last section, we will analyze the movie business. We will figure out some factors to the profit and ROI and visualize their patterns. Our goal is to give suggestions for companies and directors who want to be succesful in the movie industry.  

#### 5.1 SMART Question

##### Which genres are good selections for commercial success? By analyzing the profit and the return on investment (ROI) of different genres from 1986 - 2016 which suggestions we should give to the movie companies?

#### 5.2 Profit distribution


```{r prfnormality, include = T}

# Q-Q plot with outliers
qqnorm(movies_fam$profit, main = "Normal Q-Q plot with outliers")
qqline(movies_fam$profit, col = "red", lw = 2)

# remove outliers to have a better assumption of profit distribution
outliers <- boxplot(data=movies_fam, profit ~ genre, plot = F)$out # remove outliers
movies_fam_noOut <- subset(movies_fam, ! movies_fam$profit%in%outliers) # a subset without outliers


# Q-Q plot without outliers
qqnorm(movies_fam_noOut$profit, main = "Normal Q-Q plot without outliers")
qqline(movies_fam_noOut$profit, col = "red", lw = 2)



```

* The Q-Q plot with outliers shows a distribution with over-dispersed data due to outliers. To have a better visualization, we will remove outliers and plot the Q-Q plot again.
* The profit distribution looks different from a normal distribution. It has an extreme right skew and fatter tails than a normal distribution.
* Since the profit shows an extreme right skew, the frequency is higher to the left. If we select the mean as threshold and randomly choose a movie, this movie will likely have lower profit than the threshold. In other words, the majority of movies tend to have a lower profit than the average.
* This pattern will not statisfy movie organizations. They expect a profit distribution with less concentration on the left. A left skew distribution is an ideal result.


Now, we have a look at the 5-number summary of profit (including the mean) in table format:

```{r prfsummary, include = T}
summary(movies_fam$profit)

```

* The median is small than zero, which means that more than 50% of published movies have no profit. This shows that there is a high chance for a movie company to have a net-less with their product.
* To help movie companies to address the issue of profit, we will conduct a deeper analysis on movie profit.

#### 5.3 Profit by genres

In this part, we will identify which genre earns most profit.


```{r profitboxplots}

# plot with outliers
ggplot() +
  geom_boxplot(data = movies_fam, aes(x=genre, y=(profit*(1e-6)), fill=genre)) +
  ggtitle("Profit distribution by genres") +
  xlab("Genre") +
  ylab("Profit (millions USD)") +
  theme(plot.title = element_text(hjust = 0.5))
  
# plot without outliers
ggplot() +
  geom_boxplot(data = movies_fam_noOut, aes(x=genre, y=(profit*(1e-6)), fill=genre)) +
  ggtitle("Profit distribution by genres without outliers") +
  xlab("Genre") +
  ylab("Profit (millions USD)") +
  theme(plot.title = element_text(hjust = 0.5))
  

```

* From the first figure (with outliers), both the most net-income movie and the most net-loss movie belong to Action genre. 
* The Animation and Horror have their median lines at the highest positions and lie over zero, showing that these two genres provide good ratios of profitable movies (more than 50% of movies have profit). It seems to be a safer choice to invest on Animation and Horror rather than other genres.

* Observe Animation and Horror distribution:

```{r Ani&Hor, include = T}

# Animation normality
Ani <- subset(movies_fam_noOut, genre == "Animation") # a subset including Animation movies only
qqnorm(Ani$profit, main = "Normal Q-Q plot of Animation")
qqline(Ani$profit, col ="blue", lw = 2)
hist(Ani$profit, main ="Histogram of Animation", col = "yellow1", xlab = "Profit")

# Horror normality
Hor <- subset(movies_fam_noOut, genre == "Horror") # a subset including Horror movies only
qqnorm(Hor$profit, main = "Normal Q-Q plot of Horror")
qqline(Hor$profit, col = "blue", lw = 2)
hist(Hor$profit, main ="Historgram of Horror", col = "chocolate1", xlab = "Profit")

```

* As we see from the Q-Q plots, profit distributions of Animation and Horror look roughly normally distributed. Both distributions have a slight right skew. Although movie companies expect a left skew distribution but a slight right skew distribution is still acceptable.

* We observe the ratio of profitable movies in each genre: 

```{r ratioProfit, include = T}

gain_mv <- subset(movies_fam, movies_fam$profit > 0)
#loss_mv <- subset(movies_fam, movies_fam$profit < 0 | movies_fam$profit == 0)

genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")

ratio <- c("")
j = 1

# ratio (numbers of movie with positive income)/(numbers of movie) by each genre

for(i in genrelist) {ratio[j] <- nrow(subset(gain_mv, gain_mv$genre == i))*100/nrow(subset(movies_fam,(! (is.na(movies_fam$profit)) & (movies_fam$genre == i)))); j = j+1} # calculate ratio and save to a vector

ratio_df <- data.frame(Ratio = ratio, Genre = genrelist) # save ratio and related genre to a new dataframe
ratio_df$Ratio <- as.numeric(as.character(ratio_df$Ratio)) # change factor to numeric 
ratio_df <- ratio_df[order(ratio_df$Ratio, decreasing = T),] # ordering
ratio_df


```

* As shown in the table, Horor and Animation have highest ratio of profitable movies, matching our observation from the boxplots.
* Movie companies and directors are recommended to invest in these two categories: Animation and Horror.
* Now, we will take a look at the distribution of profit average (profit per movie) by genres.

```{r avgprofitdef}

pf = function(x) {
  genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")
  profit_avg <- c("")
  j = 1
  
  for(i in genrelist) {profit_avg[j] <- mean((subset(x, x$genre == i))$profit, na.rm = T)*(1e-3); j = j + 1}
  
  df = data.frame(meanprofit = profit_avg, genreprofit = genrelist)
  df$meanprofit <- as.numeric(as.character(df$meanprofit))

  
  p <- ggplot(df, aes(x="", y=meanprofit, fill = genreprofit)) +
    geom_bar(width=1, stat = "identity", color = "white") +
    coord_polar("y", start=0) +
    guides(fill=guide_legend(title="Genre")) +
    theme(axis.text  = element_blank(), axis.ticks = element_blank(), axis.title = element_blank()) +
    geom_text(aes(label = paste0( scales::percent(meanprofit / sum(meanprofit)))), position = position_stack(vjust = 0.5))
  

   return(p)
  
}
```


```{r pltProfit, echo=F}

# comparision of  profit per movie of each genre across different periods of time

pf(movies_new) +
  ggtitle("Profit Average distribution by Genres") +
  theme(plot.title = element_text(hjust = 0.5)) 
 
pf(movies1) +
  ggtitle("Profit Average distribution by Genres in 1986-1999") +
  theme(plot.title = element_text(hjust = 0.5)) 
 
pf(movies2) +
  ggtitle("Profit Average distribution by Genres in 2000-2009") +
  theme(plot.title = element_text(hjust = 0.5)) 
 
pf(movies3) +
  ggtitle("Profit Average distribution by Genres in 2010-2016") +
  theme(plot.title = element_text(hjust = 0.5)) 
  
  


```

* As we see from our pie charts, Animation, Adventure and Horor are always the most profitable genres. 
* Animation is the genre giving most profit per movie. It is understandable since Animation does not require budget to hire actors and investment on visual effects and movie set.
* Adventure and Horror places next since those genres also do not require a lot of investment but achieve good income.
* Although Action is the one of the genres with most votes, its movies do not earn good profit. The reason is that Action movies require a considerable investment on actors, visual effects and movie set which cost a lot of expenditure.


Calculate profit and mapping plot (x=budget, y=profit, color by genre/rating)

```{r mappingProfit, echo=F, include = T, warning = F}
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=(budget*(1e-6)), y=(profit*(1e-6)), col = genre)) +
  xlab("Budget (millions USD)") +
  ylab("Profit (millions USD)") +
  ggtitle("Profit vs Budget map") +
  theme(plot.title = element_text(hjust = 0.5)) 

```

* As we see that, low budget movies tend to make profit, while higher budget movies are at risk of having losses: As budget value increases, the range of gross-loss value also increases.
* High-profit movies (mostly action genre) require pretty high budgets (more than `r quantile(movies_fam$budget, 0.95, na.rm=T)*(1e-6)` m$, i.e 95 percentile value).
* Horror seems to have a low budget but their income is impressive. Horror seems to be the most successful movie genre. We will confirm this with analysis on Return on Investment (ROI).


#### 5.4 Return on Investment

* Return on Investment is calculated as the ratio of Profit over Budget: Return on Investment (ROI) = (Profit) / (Budget)
* Return on Investment determines whether a movie is successful.


```{r mappingROI, echo=F, include = T, warning = F}

# plot return on investment ~ budget
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=(budget*(1e-6)), y=roi, col = genre)) +
  xlab("Budget (millions USD)") +
  ylab("Return on Investment (%)") +
  ggtitle("Return on Investment vs Budget map") +
  theme(plot.title = element_text(hjust = 0.5)) 



```

* As we claimed, Horror movies have low budget and high profit, resulting in a very high return on investment (ROI=Profit/Budget). 
* We can conclude that Horror is the most successful genre in movie business.
* However, it is hard for us to visualize other genre. We need to remove outliers for a better observation.
* We remove outliers and create new plots:


```{r remappingROI, echo=F, include = T, warning = F}

# plot return on investment ~ budget wihout outliers
ggplot() +
  geom_point(data=subset(movies_fam,! movies_fam$roi%in%boxplot(roi ~ genre, data=movies_fam, plot = F)$out), mapping=aes(x=(budget*(1e-6)), y=roi, col = genre)) +
  xlab("Budget (millions USD)") +
  ylab("Return on Investment (%)") +
  ggtitle(("Return on Investment vs Budget map without outliers")) +
  theme(plot.title = element_text(hjust = 0.5)) 


# boxplot ROI by genres without outliers
ggplot(data=subset(movies_fam,! movies_fam$roi%in%boxplot(roi ~ genre, data=movies_fam, plot = F)$out), aes(x=genre, y=roi, fill=genre)) +
  geom_boxplot() +
  xlab("Genre") +
  ylab("Return on Investment (%)") +
  ggtitle(("Return on Investment distribution by genres without outliers")) +
  theme(plot.title = element_text(hjust = 0.5)) 

  

```

* As expected, Horror movies still possess high return on investment. 
* Action movies seem not to do well in Return on Investment. Due to its characteristics, Action requires high expenditures which lead to its low return on investment.
* The Return on Investment tends to reduce as the budget increases. Therefore, low budget movie tend to be more successful than high budget movie in movie commerce.

#### 5.5 Profit and Votes Relationship

```{r ProfitvsVotes, echo=F, include = T, warning = F}
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=votes, y=(profit*(1e-6)), col = genre)) +
  xlab("Votes") +
  ylab("Profit (millions USD)") +
  ggtitle("Profit vs Votes map") +
  theme(plot.title = element_text(hjust = 0.5)) 

```

* There is no clear relationship between profit and votes. Votes seem not to be a factor to profit.

* The only thing we see is that movies with more than 500,000 votes seem to have a safe income, most of these movies have positive profits. The major of movies with net-losses belong to the group of below-500,000 votes movies.

#### 5.6 Profit and Scores Relationship

```{r ProfitvsScore, echo=F, include = T, warning = F}
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=score, y=(profit*(1e-6)), col = genre)) +
  xlab("Score") +
  ylab("Profit (millions USD)") +
  ggtitle("Profit vs Score map") +
  geom_vline(linetype="dashed", aes(xintercept = 3.5)) +
  geom_hline(linetype="dashed", aes(yintercept = 350)) +
  theme(plot.title = element_text(hjust = 0.5)) 

```


* Score seems not to contribute a lot to profit. There are only two remarkle features: Movies that earn very high profits (more than 350 m$) often achieve good scores (around 7.0), whereas very low score movies (less than 3.5) earn small profits and can also have net-losses.

* Except movies with significantly high profits or significantly low scores, the others have their profits vary widely and unpredictably in every score grade.


### Chapter 6: Conclusion

* In summary, there are many insights from our analysis on movie industry:
1. USA is the biggest movie industry since around 74% of published movies worldwide come from this country. However, USA is not an appropriate sample to study the global movie industry because there is no evidence that USA data match the wolrd data.
2. From our testing, we observe a relationship between movie genre and ratings. In other words, genre is a factor to decide a movie rating. It is understandable since different categories are determined by their target audiences, contents and level of violence which are also the basis to restrict the range of viewers.
3. Although there is no clear relationship between movie scores and its votes, we can conclude that a movie with low score will always receive few votes and a movie with lots of votes will score well.
4. We do not have enough evidence to determine if movie genre has any effect on customers' score and votes.
5. The most profitable movie belongs to Action but the most successful genre is Horror. Animation has highest profit per movie and maintain a stability in making income over the three decades. 

* Finally, we have a suggestions for movie companies and directors:
1. If companies have little budget or want to navigate to the success of their products, they should invest on Horror movies. Horror also possess a decent profit per movie. 
2. If a company prioritizes the profit, Animation is a safe and stable choice.
3. If a company has strong economic potential and possesses expert directors and actors, they can invest on Action genre. Although this selection is risky but it can reward a considerable amount of income if the movie is successful.
