---
title: "Project 1 - DATS 6101"
author: "Randomly-Generated: Le Tran Hieu "
date: "9/29/2019"
output:
   html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r define_loadPkg, echo=F}
loadPkg = function(x) { if (!require(x,character.only=T, quietly =F)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T, quietly=T)) stop("Package not found") } }
```

### Chapter 1: Introduction

We have seven chapters in this project based on a dataset of movies. Chapter one and seven are introduciton and conclusion. We have five seperate SMART questions in total. And each chapter from two to six contains a main question and several sections that help us to analyze the main question.

For the SMART questions, for example, we wonder if the audiences are more interested in any specific movie genre from late 20th century till now. Also, does a movie with more people rating it have higher score? Based on the current commercial movie market, is there a movie genre that allows the movie company to achieve commercial success relatively easier. Moreover, does the country with the most number of movie pubished represent the movie industry? We use a dataset including 6820 movies from 1986-2016 to analyze the question we raised.

Firstly, we have to clean the data by filtering the columns we are going to analyze and dealing with some special cases like missing datas. For each question analyze, we include data visualization and statistical tests. The answer for our questions are basically generated by the tests we apply on our dataset and comparison of several variables we do according to the plots. 



### Chapter 2: Description of Data

#### 2.1 Source Data

The links for the dataset is: https://www.kaggle.com/danielgrijalvas/movies , which was separated from IMDB.
 
#### 2.2 Data Explanation

There are 6820 movies in the dataset from 1986-2016. 

Explanation for columns attributes:

* budget: the budget of a movie
* company: the movie company
* country: the country that produces the movie
* director: the director of the movie
* genre: the main genre of the movie
* gross: revenue of the movie
* name: name of the movie
* rating: a system determines the suitability of movies for different audiences (classified by ages)
* released: release date (YYYY-MM-DD)
* runtime: the duration of the movie
* score: IMDB user rating
* votes: the number of user votes
* star: the main actor/actress
* writer: the writer of the movie
* year: the year of release



For our project, we will remove all movies that are television or DvDs versions (which did not follow the Motion Picture Association of America (MPAA) film rating system ) and focus on movies published at theatre/cinema since our goal is to research the theatrical exhibition.

#### 2.3 Data Cleaning

Step 1: Load dataset from csv file and save it as movies dataframe.


```{r load_data, echo = F, message = F}

loadPkg("ggplot2")
loadPkg("dplyr")

# read dataset
movies <- read.csv("movies.csv")
str(movies)

```

Step 2: Clean the dataframe. We will save new dataset as movies_new

* Change all movies with budget = 0 to NA.

```{r NAs, echo = F}

# assign NA to movies with budget = 0 
movies$budget <- replace(movies$budget,movies$budget==0, NA)

```

* Remove movies that are TV or DVDs versions.

```{r ratings_rework, echo = F}

# filter for movies with MPAA film rating system 
movies_new <- subset(movies, movies$rating%in%c("G","NC-17","PG-13","R","PG"))


# apply factor again to new dataset
movies_new$rating <- factor(movies_new$rating) 
movies_new$genre <- factor(movies_new$genre)
```

* Remove duplicated data. Fortunately, all the data are distinct.

```{r duplicates, echo = F}
# remove duplicates if they exist
movies_new <- unique(movies_new)

```

* remove columns that will not be used

```{r rmcol, echo = F}

movies_new <- subset(movies_new, select = - c(company,director,released,runtime,star,writer))

```

* Since we will work on profit and return on investment later, we will create 2 new columns in the dataframe

```{r createcol, echo = F}

movies_new$profit <- movies_new$gross - movies_new$budget # create profit column

movies_new$roi <- (movies_new$profit)*100/(movies_new$budget) # create ROI column



```

* Show our new dataframe movies_new

```{r datastructure, echo = F}
# show dataframe structure
str(movies_new)


```

#### 2.3 Data Visualization

```{r datahist, echo = F, include = T}

ggplot(data = movies_new, aes(x = year)) +
  geom_histogram(alpha = 0.5, fill = "blue", col = "red", binwidth = 1) +
  ggtitle("Histogram of published movies from 1986-2016") +
  xlab("Year") +
  ylab("Number of Movies") +
  theme(plot.title = element_text(hjust=0.5))
  


```

### Chapter 3: Movie categories
#### 3.1 SMART Question
##### Does the genre has any effects on the classification of audiences and which genre achieve highest voting averages in three time intervals: 1986-1999, 2000-2010, and 2010-2016?

#### 3.2 Number of published movies by genres:

* We will take a look at the number of published movies in each genre from 1986-2016. A

```{r count, echo = F, include = T}

# show the number of published movies by genre
table(movies_new$genre)

# save new dataset with famous genres
movies_fam <- subset(movies_new, movies_new$genre%in%c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror"))
movies_fam$genre <- factor(movies_fam$genre)

```


```{r pltcount, echo = F, include =T}

ggplot(movies_new, aes(x=genre, fill = genre)) +
  geom_bar(stat="count", ) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("The number of movies published by genre from 1986 to 2016") +
  guides(fill = guide_legend(title="Genre"))


```

* As we see from the table and the bar charts, there were 8 famous genres with many published movies: Action, Adventure, Animation, Biography, Comedy, Crime, Drama, and Horror, in which Action, Comedy and Drama are the genres that have most products, whereas Family, Fantasy, Musical, Mystery, Romance, Sci-Fi, Thriller, War and Western movies seemed not to be preferred by film companies and directors. Indeed, 8 famous genres also included the elements of Family, Fantasy, Musical, Mystery, Romance, Sci-Fi, Thriller. And War and Western movies seem not to interest viewers as they focus a lot on historical events and stories which appear frequently in papers, books and documentaries.

* Later, any small size samples (the number of movies published is smaller than 50) will not be inlcuded in analysis to ensure reasonable results. We will only focus on most famous genres.

```{r save_a_genrelist, echo = F, inlcude = F}
# save 8 most famous genres into a vector
fullgenrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Family","Fantasy","Horror","Musical","Mystery","Romance", "Sci-Fi","Thriller", "War","Western")
genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")


```


#### 3.3 Test the dependence of genre and rating:

Movie ratings explanation:

* G – General Audiences: All ages admitted. Nothing that would offend parents for viewing by children.
* PG – Parental Guidance Suggested: Some material may not be suitable for children. Parents urged to give "parental guidance". May contain some material parents might not like for their young children.
* PG-13 – Parents Strongly Cautioned: Some material may be inappropriate for children under 13. Parents are urged to be cautious. Some material may be inappropriate for pre-teenagers.
* R – Restricted: Under 17 requires accompanying parent or adult guardian. Contains some adult material. Parents are urged to learn more about the film before taking their young children with them.
* NC-17 – Adults Only: No One 17 and Under Admitted. Clearly adult. Children are not admitted.

* Create Contigency Table:

``` {r contable, echo = F, include = T}

# create contigency table

contable1 = table(movies_new$rating, movies_new$genre)
contable1

# create contigency table
#contable2 = table(movies_fam$rating, movies_fam$genre)
#contable2

````

* Chi-squared Test:

```{r chitest, echo=F, warning=FALSE, include=T}

# chi-square test
chitest1 <- chisq.test(contable1)

# results
chitest1
chitest1$statistic
chitest1$parameter
chitest1$expected
chitest1$p.value


# chi-square test with famous genres (large-size samples)


# chi-square test
#chitest2 <- chisq.test(contable2)

# results
#chitest2
#chitest2$statistic
#chitest2$parameter
#chitest2$expected
#chitest2$p.value

```

* Null Hypothesis: The movie genre and its rating are independent.
* The p-value in the chi-squared test is `r chitest1$p.value` that is much smaller than the significance level (0.05). Therefore, we reject the null.
* Conclusion: The movie genre and rating are NOT independent.
* Therefore, movie genre and rating are related to each other.

#### 3.4 Genres with most votes

* To have a better observation, we will divide our reseacrh period into three time intervals: 1986-1999, 2000-2009, 2010-2016
* Then, we save our movies data into three datasets.

``` {r databyperoid, echo= F}
movies1 <- subset(movies_fam, movies_fam$year%in%c(1986:1999)) # dataset 1 from 1986-1999
movies2 <- subset(movies_fam, movies_fam$year%in%c(2000:2009)) # dataset 2 from 2000-2009
movies3 <- subset(movies_fam, movies_fam$year%in%c(2010:2016)) # dataset 3 from 2010-2016
#str(movies1)
#str(movies2)
#str(movies3)
```

```{r def, echo = F}
pltm = function(x) {
  genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")
  vote_avg <- c("")
  j = 1
  
  for(i in genrelist) {vote_avg[j] <- mean((subset(x, x$genre == i))$votes, na.rm = T)*(1e-3); j = j + 1}
  
  df = data.frame(meanvote = vote_avg, genrevote = genrelist)
  df$meanvote <- as.numeric(as.character(df$meanvote))
  
  p <- ggplot(df, aes(x="", y=meanvote, fill = genrevote)) +
    geom_bar(width=1, stat = "identity", color = "white") +
    coord_polar("y", start=0) +
    guides(fill=guide_legend(title="Genre")) +
    theme(axis.text  = element_blank(), axis.ticks = element_blank(), axis.title = element_blank()) +
    geom_text(aes(label = paste0( scales::percent(meanvote / sum(meanvote)))), position = position_stack(vjust = 0.5))
  

  
  
  return(p)


}
```




```{r genrevotes, echo = F}

# Mean votes distribution 1986-1999
pltm(movies1) +
  ggtitle("Vote Means distribution by Genres in 1986-1999") +
  theme(plot.title = element_text(hjust = 0.5)) 
 

# Mean votes distribution 2000-2009
pltm(movies2) +
  ggtitle("Vote Means distribution by Genres in 2000-2009") +
  theme(plot.title = element_text(hjust = 0.5))

# Mean votes distribution 2010-2016
pltm(movies3) +
  ggtitle("Vote Means distribution by Genres in 2010-2016") +
  theme(plot.title = element_text(hjust = 0.5))


```

* As we see from the plots: Action, Animation, Adventure and Crime are always the top genres with most votes.
* During the first period, Animation and Crime were the most favorite genres. In the second period, Action and Adventure grew as the most favorite genres and in recent years, these two genres still have maintained their posistions. 
* In contrast, Animation and Crime had a significance backward after three periods. Although they still are in top 4 genres with most votes, their first and second places were taken by Action and Adventure.
* Comedy and Horror seems to be the least interested genres as they always receive least votes.


### Chapter 4: Movie Scores
#### 4.1 SMART Question
##### Is there any relationships between the votes and the scores and are the score means the same across all genres and all ratings?


#### 4.2 Scores and Votings Relationship

* We divide votes and scores into different categories:

Votes:

- High votes: from 1,000,000  
- Medium votes: from 500,000 to below 1,000,000
- Low votes: below 500,000

Scores:

- Good score: from 7.5
- Medium score: from 5.0 to below 7.5
- Bad score: below 5.0

Now, we plot the distribution of votings across the scores using boxplots and mapping plots"

```{r pltVotesVsScore, echo = F, include = T, message=F}

high_votes <- subset(movies_new, votes > 1e+6 | votes == 1e+6)
medium_votes <- subset(movies_new, votes < 1e+6 & votes > 500000 | votes == 500000)
low_votes <- subset(movies_new, votes < 500000)



ggplot() +
  ggtitle("Score distribution with different groups of votes") +
  geom_boxplot(data=high_votes, aes(x=1,y=score,fill="1")) +
  geom_boxplot(data=medium_votes, aes(x=2,y=score,fill="2")) +
  geom_boxplot(data=low_votes, aes(x=3,y=score,fill="3")) +
  theme(plot.title = element_text(hjust=0.5),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill=guide_legend(title = "Vote categories")) +
  scale_fill_manual(labels=c("High","Medium","Low"),values = c("blue","red","yellow"))
  
ggplot() +
  ggtitle("Votes distribution over score") +
  geom_point(data=movies_new,mapping=aes(x=score,y=votes,col=rating)) +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_hline(linetype="dashed", aes(yintercept = 1e+6)) + 
  geom_vline(linetype="dashed", aes(xintercept = 7.5)) +
  geom_hline(linetype="dashed", aes(yintercept = 500000)) + 
  geom_vline(linetype="dashed", aes(xintercept = 5.0)) +
  annotate("text", x = 8.6, y = 15*(1e+5), label = "High Votes\nHigh Score") +
  #annotate("text", x = 6.0, y = 15*(1e+5), label = "High Votes\nMedium Score") +
  annotate("text", x = 2.5, y = 15*(1e+5), label = "High Votes\nLow Score") +
  #annotate("text", x = 8.6, y = 750000, label = "Medium Votes\nHigh Score") +
  #annotate("text", x = 6.0, y = 750000, label = "Medium Votes\nMedium Score") +
  #annotate("text", x = 2.5, y = 750000, label = "Medium Votes\nLow Score") +
  annotate("text", x = 8.6, y = 250000, label = "Low Votes\nHigh Score") +
  #annotate("text", x = 6.0, y = 250000, label = "Low Votes\nMedium Score") +
  annotate("text", x = 2.5, y = 250000, label = "Low Votes\nLow Score")

#qqnorm(movies_new$votes)
#qqnorm(movies_new$score)

```

* As shown in boxplots, the movies with highest votes (above 1 million votes) tend to have high scores and the movies with low scores (below 5) tend to get bad votes. However, high score movies do not certainly have high votes and low votes movies do not mean that they have low score. 
* The mapping plot also gives us the same observation, every movies with votes over 1 million always possesses a high score. And at around 500000 votes and below, the score varies extensively. Movies with scores below 5.0 have low votes.
* This illustrates that people often highly grade their favorite movies (with high votes). Meanwhile, they do not recommend bad movies (low score). The reason for the large variety of scores at medium votes ( below 500000) may be due to the genre of the movies that affect the number of audiences. Many genres have good quality of movies but they are not of interest to audiences such as Biography, War, Scifi, Drama, which makes them get low votes even they are highly graded.


#### 4.3 Score means of different genres and different ratings

##### By genres:

* Anova test score means by famous genres: (we will not include genres with low sample sizes < 50)

```{r AOV_score&genre, echo = F, include =T}

anova_sg <- aov(score ~ genre, movies_fam)
summary(anova_sg)

```

* Null Hypothesis: The score means of all genres are the same.
* The p-value of the test is `r summary(anova_sg)[[1]][["Pr(>F)"]][1]`, which is very close to 0 and much smaller than significance level (0.05).
* Therefore, we reject the null and conclude that the scores means of different genres are not all the same.

```{r pltScore&Genre, echo = F, include = T}
ggplot() +
  geom_boxplot(data=movies_fam, aes(x=genre, y=score, fill=genre)) +
  ggtitle("Score distribution by genres") +
  theme(plot.title=element_text(hjust=0.5))
  
```

* As shown in the boxplots, the following pairs have pretty the same tall boxplots and overlapped interquartile ranges: Action-Comedy, Animation-Crime, Animation-Drama and Crime-Drama.
* This illustrates that these pairs seem to be NOT significant in their mean values. 
* The other pairs show pretty clear differences in their boxplots, meaning that they are significant at 0.05 level.
* We will conduct an ad hoc procedure using Tukey test to see which pairs are significant and check whether our visualization is correct. 

```{r TukeyTest1, echo = F, include = T}

tukey1 <- TukeyHSD(anova_sg)
tukey1
```


* As we see from our Tukey test, there are no significant differences in score means of the following pairs: Action-Comedy, Animation-Crime, Animation-Drama, Crime-Drama, while the other pairs are significant.
* Our observation is pretty correct.

##### By ratings:


```{r AOV_score&rating}

anova_sr <- aov(score ~ rating, movies_fam )
summary(anova_sr)

```

* Null Hypothesis: The score means of all ratings are the same.
* The p-value of the test is `r summary(anova_sr)[[1]][["Pr(>F)"]][1]`, which is very close to 0 and much smaller than significance level (0.05).
* Therefore, we reject the null and conclude that the scores means of different ratings are not all the same.

```{r pltScore&Rating, echo = F, include = T}
ggplot() +
  geom_boxplot(data=movies_fam, aes(x=rating, y=score, fill=rating)) +
  ggtitle("Score distribution by ratings") +
  theme(plot.title=element_text(hjust=0.5))
  
```

* As shown in the boxplots, the five boxes overlapped with each other and their medians also line within other boxes' interquartiles. Therefore, if we observe these boxplots, we will feel that their means are not significant.
* We will conduct a Tukey Test to see whether our visualization is correct.

```{r TukeyTest2, echo = F, include = T}

tukey2 <- TukeyHSD(anova_sr)
tukey2
```

* As we see from our Tukey test, there are significant differences in score means of the following pairs: PG-13 & R, PG & R, PG & PG-13, G & PG.
* The other pairs are not significnace at 0.05 level.
* Our obsevation from the boxplots is not correct at all.


### Chapter 5: Movie Commerce

#### 5.1 SMART Question

##### Which genres are good selections for commercial success? Analyzing the profit and the return on investment (ROI) of different genres from 1986 - 2016 to give suggestions to the movie companies. 

#### 5.2 Profit



```{r profitboxplots, echo=F, warning=F}

outliers <- boxplot(data=movies_fam, profit ~ genre, plot = F)$out

movies_fam_noOut <- subset(movies_fam, ! movies_fam$profit%in%outliers)

ggplot() +
  geom_boxplot(data = movies_fam, aes(x=genre, y=(profit*(1e-6)), fill=genre)) +
  ggtitle("Profit distribution by genres") +
  xlab("Genre") +
  ylab("Profit (millions USD)") +
  theme(plot.title = element_text(hjust = 0.5))
  

ggplot() +
  geom_boxplot(data = movies_fam_noOut, aes(x=genre, y=(profit*(1e-6)), fill=genre)) +
  ggtitle("Profit distribution by genres without outliers") +
  xlab("Genre") +
  ylab("Profit (millions USD)") +
  theme(plot.title = element_text(hjust = 0.5))
  

```

* Both the most net-income movie and the most net-loss movie belong to Action genre. 
* The Animation and Horror have their median lines at the highest positions, showing that these two genres provide good ratio of profitable movies. It is a safe choice for companies and directors to produce movies of Animation and Horor.
* Animation is significantly postive-skewed, meaning that its mean value is very high compare to its median. We can predict that Animation has the highest average of profit (estimated profit per film/movie).

* We observe the ratio of profitable movies in each genre: 


```{r ratioProfit, echo=F, include=T}

gain_mv <- subset(movies_fam, movies_fam$profit > 0)
#loss_mv <- subset(movies_fam, movies_fam$profit < 0 | movies_fam$profit == 0)

genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")

ratio <- c("")
j= 1

# ratio (numbers of movie with positive income)/(numbers of movie) by each genre

for(i in genrelist) {ratio[j] <- nrow(subset(gain_mv, gain_mv$genre == i))*100/nrow(subset(movies_fam,(! (is.na(movies_fam$profit)) & (movies_fam$genre == i)))); j = j+1} # calculate ratio and save to a vector

ratio_df <- data.frame(Ratio = ratio, Genre = genrelist) # save ratio and related genre to a new dataframe
ratio_df$Ratio <- as.numeric(as.character(ratio_df$Ratio)) # change factor to numeric 
ratio_df <- ratio_df[order(ratio_df$Ratio, decreasing = T),] # ordering
ratio_df


```

* As shown in the table, Horor and Animation have highest ratio of profitable movies, matching our observation from the boxplots.
* Now, we will take a look at the distribution of profit average (profit per movie) by genres

```{r avgprofitdef, echo = F, include = F}

pf = function(x) {
  genrelist = c("Action","Adventure","Animation","Biography","Comedy","Crime","Drama","Horror")
  profit_avg <- c("")
  j = 1
  
  for(i in genrelist) {profit_avg[j] <- mean((subset(x, x$genre == i))$profit, na.rm = T)*(1e-3); j = j + 1}
  
  df = data.frame(meanprofit = profit_avg, genreprofit = genrelist)
  df$meanprofit <- as.numeric(as.character(df$meanprofit))
  
  p <- ggplot(df, aes(x="", y=meanprofit, fill = genreprofit)) +
    geom_bar(width=1, stat = "identity", color = "white") +
    coord_polar("y", start=0) +
    guides(fill=guide_legend(title="Genre")) +
    theme(axis.text  = element_blank(), axis.ticks = element_blank(), axis.title = element_blank()) +
    geom_text(aes(label = paste0( scales::percent(meanprofit / sum(meanprofit)))), position = position_stack(vjust = 0.5))
  

   return(p)
  
}
```


```{r pltProfit, echo=F}

# comparision of  profit per movie of each genre across different periods of time

pf(movies_new) +
  ggtitle("Profit Average distribution by Genres") +
  theme(plot.title = element_text(hjust = 0.5)) 
 
pf(movies1) +
  ggtitle("Profit Average distribution by Genres in 1986-1999") +
  theme(plot.title = element_text(hjust = 0.5)) 
 
pf(movies2) +
  ggtitle("Profit Average distribution by Genres in 2000-2009") +
  theme(plot.title = element_text(hjust = 0.5)) 
 
pf(movies3) +
  ggtitle("Profit Average distribution by Genres in 2010-2016") +
  theme(plot.title = element_text(hjust = 0.5)) 
  
  


```

* As we see from our pie charts, Animation is the genre giving most profit per movie. It is understandable since Animation does not require budget to hire actors and investment on visual effects and movie set.
* Adventure and Horror places next since those genres also do not require a lot of investment but achieve good income.
* Although Action is the one of the genres with most votes, it do not earn good profit. The reason is that Action movies require a considerable investment on actors, visual effects and movie set which cost a lot of expenditure.


Calculate profit and mapping plot (x=budget, y=profit, color by genre/rating)

```{r mappingProfit, echo=F, include = T, warning = F}
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=(budget*(1e-6)), y=(profit*(1e-6)), col = genre)) +
  xlab("Budget (millions USD)") +
  ylab("Profit (millions USD)") +
  ggtitle("Profit vs Budget map") +
  theme(plot.title = element_text(hjust = 0.5)) 

```

* As we see that, low budget movies tend to make profit, while higher budget movies are at risk of having losses: As budget value increases, the range of gross-loss value also increases.
* High-profit movies (mostly action genre) require pretty high budgets (more than `r quantile(movies_fam$budget, 0.95, na.rm=T)*(1e-6)` m$, i.e 95 percentile value).
* Horror seems to have a low budget but their income is impressive, which will be reflected more clearly by Return on Investment plots.


#### 5.3 Return on Investment

* Return on Investment is calculated as the ratio of Profit over Budget: Return on Investment (ROI) = (Profit) / (Budget)
* Return on Investment determines whether a movie is successful.




```{r mappingROI, echo=F, include = T, warning = F}

# plot return on investment ~ budget
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=(budget*(1e-6)), y=roi, col = genre)) +
  xlab("Budget (millions USD)") +
  ylab("Return on Investment (%)") +
  ggtitle("Return on Investment vs Budget map") +
  theme(plot.title = element_text(hjust = 0.5)) 



```

* As we claimed, Horror movies have low budget and high profit, resulting in a very high return on investment (ROI=Profit/Budget). However, it is hard for us to visualize other values. We need to remove outliers for a better observation.

* We remove outliers and create new plots:


```{r remappingROI, echo=F, include = T, warning = F}

# plot return on investment ~ budget wihout outliers
ggplot() +
  geom_point(data=subset(movies_fam,! movies_fam$roi%in%boxplot(roi ~ genre, data=movies_fam, plot = F)$out), mapping=aes(x=(budget*(1e-6)), y=roi, col = genre)) +
  xlab("Budget (millions USD)") +
  ylab("Return on Investment (%)") +
  ggtitle(("Return on Investment vs Budget map without outliers")) +
  theme(plot.title = element_text(hjust = 0.5)) 


# boxplot ROI by genres without outliers
ggplot(data=subset(movies_fam,! movies_fam$roi%in%boxplot(roi ~ genre, data=movies_fam, plot = F)$out), aes(x=genre, y=roi, fill=genre)) +
  geom_boxplot() +
  xlab("Genre") +
  ylab("Return on Investment (%)") +
  ggtitle(("Return on Investment distribution by genres without outliers")) +
  theme(plot.title = element_text(hjust = 0.5)) 

  

```

* As expected, Horror movies still possess high return on investment. Action movies which often required high expenditures have low return on investment despite their high incomes.

* The Return on Investment reduces as the budget increases.

#### 5.4 Profit and Votes Relationship

```{r ProfitvsVotes, echo=F, include = T, warning = F}
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=votes, y=(profit*(1e-6)), col = genre)) +
  xlab("Votes") +
  ylab("Profit (millions USD)") +
  ggtitle("Profit vs Votes map") +
  theme(plot.title = element_text(hjust = 0.5)) 

```

* There is no clear relationship between profit and votes. Votes seem not to be a factor to profit.

* The only thing we see is that movies with more than 500,000 votes seem to have a safe income, most of these movies have positive profits. The major of movies with net-losses belong to the group of below-500,000 votes movies.

#### 5.5 Profit and Scores Relationship

```{r ProfitvsScore, echo=F, include = T, warning = F}
ggplot() +
  geom_point(data=movies_fam, mapping=aes(x=score, y=(profit*(1e-6)), col = genre)) +
  xlab("Score") +
  ylab("Profit (millions USD)") +
  ggtitle("Profit vs Score map") +
  geom_vline(linetype="dashed", aes(xintercept = 3.5)) +
  geom_hline(linetype="dashed", aes(yintercept = 350)) +
  theme(plot.title = element_text(hjust = 0.5)) 

```


* Score seems not to contribute a lot to profit. There are only two remarkle features: Movies that earn very high profits (more than 350 m$) often achieve good scores (around 7.0), whereas very low score movies (less than 3.5) earn small profits and can also have net-losses.

* Except movies with significantly high profits or significantly low scores, the others have their profits vary widely and unpredictably in every score grade.


### Chapter 6: Movie industry in different countries
#### 6.1 SMART Question
##### Which country has the most number of published films and movies over 30 years and does this country can represent for the whole dataset in reference to movie genres?

#### 6.2 Country with the highest number of movies

```{r country, echo = F}

# Name the country with max counts

name_max <- names(which.max(table(movies_new$country)))

# percentage that country accounts for

pp_max <- nrow(subset(movies_new, movies_new$country == name_max))*100 / nrow(movies_new)

```

* The country with the most number of published movies over 30 years is `r name_max`.
* `r name_max` accounts for approximately `r round(pp_max)` percentage of total number of movies.


#### 6.3 Test if the country can represent the whole dataset

```{r GoodnessOfFit, echo = F, include = T}

USA_movies <- subset(movies_fam, movies_fam$country == "USA")

exp_freq <- c("")
USA_freq <- c("")

# calculate expected frequencies
j = 1
for(i in genrelist) {exp_freq[j] <- nrow(subset(movies_fam, movies_fam$genre == i))/ nrow(movies_fam); j = j + 1}
exp_freq <- as.numeric(as.character(exp_freq))

# calculate observed frequencies
k = 1
for(i in genrelist) {USA_freq[k] <- nrow(subset(USA_movies, USA_movies$genre == i)); k = k + 1}
USA_freq <- as.numeric(as.character(USA_freq))

# Chi-test for Goodness of Fit
chiGoF <- chisq.test(USA_freq, p = exp_freq)


# results
chiGoF

 

```

* Null Hypothesis: The USA data distribution matches the whole dataset.
* The p-value of the test is `r chiGoF$p.value` which is much smaller than significance level 0.05.
* We reject the null, the USA data distribution does not match the whole dataset at 0.05 level.


### Chapter 7: Conclusion